The role of advection versus diffusion in intraventricular transport

### 2025-02-05 paper/main.tex
caption figure 2

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask: what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the \fixme{different-characteristics}
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \mer{MER: I think you should show on the same color scale here in a)
    (the range of values lets you and then it is possible to compare).
    Split a into a, b, c. Show e) first (so that e) becomes a)}.
    \textbf{a.} Slices in the $xz$-plane ($y=0.140$ mm,
    the center of the geometry) showing the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for three
    time instants $t=45.0$ s, $t=112.6$ s, $t=450.5$ s.
    Note that the colorbars indicating the magnitude of $c$ have different limits for each column.
    \textbf{b--d.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively.
    \textbf{e.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{f.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

>>>>

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask: what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the \fixme{different-characteristics}
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slice of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for eight time instants covering the
    time range from 45 seconds to 790 seconds, corresponding to 100 to 1750 cardiac cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    condentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}
